---
title: "Global Disease Burden Analysis"
author: "Ijeoma Iwuora"
output: github_document
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

This document compares the Disability-Adjusted Life Year (DALY) diseases burden for three countries: Benin, Cameroon, and Nigeira. We compare the burden due to communicable, maternal, neonatal, and nutritional diseases (CMNN), non-communicable diseases (NCDs), and overall disease burden for these countries. 

The source data is from the Institute for Health Metrics and Evaluation (IHME) Global Burden of Disease (GBD) study. 



# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here,janitor)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))


```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
ncd_estimates <- data_ncd %>% 
filter(Entity %in% c("Benin", "Nigeria", "Cameroon")) %>% 
  clean_names()

# You should pivot the data to show each country in a separate column.
ncd_estimate_wide <- ncd_estimates %>% 
  pivot_wider(names_from = entity, values_from = dal_ys_disability_adjusted_life_years_non_communicable_diseases_sex_both_age_age_standardized_rate)


# Use kable() from the knitr package to render the table.
kable(ncd_estimate_wide)
```

## Plot Showing Trends in NCD Burden Over Time 

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

ggplot(ncd_estimates, aes(x = year, y = dal_ys_disability_adjusted_life_years_non_communicable_diseases_sex_both_age_age_standardized_rate, color = entity)) +
  geom_line() +
  labs(title = "Trend of NCD DALY burden for Benin, Nigeria, and Cameroon",
       x = "Year",
       y = "DALYs") +
  theme_minimal() +
  scale_color_manual(values = c("Benin" = "blue", "Nigeria" = "red", "Cameroon" = "green"))  
```

## Summary of NCD Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.

## Nigeria and Benin have experienced a downward trend since the 1990s, while Cameroon exhibited an upward trend between 1990 and the early 2000s before experiencing a decline. Despite this, Nigeria has the lowest burden of NCD DALYs among the three countries.


