# Overall Disease Burden

```{r message = FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here, janitor)

# Import the overall disease data
data_overall <- read_csv(here("data", "burden-of-disease-all-causes.csv"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r warning = FALSE}
# Here render a table for the DALY burden over time for the three countries
subdata_ovrl <- data_overall %>% 
  clean_names() %>% 
  rename(dalys = dal_ys_disability_adjusted_life_years_all_causes_sex_both_age_age_standardized_rate) %>% 
  filter(code %in% c("BEN","CMR", "NGA"))

# You should pivot the data to show each country in a separate column.
subdata_ovrl_wide <- subdata_ovrl %>%  
  pivot_wider(
    id_cols = year,
    names_from = entity,
    values_from = dalys
  )

# Use kable() from the knitr package to render the table.
<<<<<<< Updated upstream
data_overall_clean <- data_overall %>% filter(Entity %in% c("Benin", "Nigeria", "Cameroon")) %>% clean_names() %>% 
  rename(DALYS = "dal_ys_disability_adjusted_life_years_all_causes_sex_both_age_age_standardized_rate")

overall_wide <- data_overall_clean %>%  pivot_wider(names_from = entity, 
                             values_from = DALYS, year) 
kable(overall_wide)
=======
kable(subdata_ovrl_wide)
>>>>>>> Stashed changes

```

## Plot Showing Trends in Overall Disease Burden Over Time

<<<<<<< Updated upstream
```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.


overall_long <-  overall_wide %>% pivot_longer(names_to = "country",
                                       values_to = "value", -year)

overall_long %>% ggplot(aes( y= value, x= year, color= country))+
  geom_line()+
  labs(y= "DALY")+
  labs(title = "Trends in overall DALY burden over time by country")
=======
```{r message = FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color

subdata_ovrl %>% 
  ggplot(aes(x = year, y = dalys, group = entity, color = entity)) +
    geom_line()

>>>>>>> Stashed changes
```

## Summary of Overall Disease Burden Findings

Overall there was a decline in the overall burden DALY diseases across all 3 countries.

